Viscous Erosion in Porous Media

### 2018-03-25 docs/methods/erosion_methods.tex
NM

	As the body shrinks in accordance with \eqref{area_predict}, the drag that it exerts decreases at a related rate, which offers a second check on our numerics. In Fig.~\ref{area_drag}(b), we show the horizontal drag on the simulated body as it erodes (black curve). The drag, $F_D$, is calculated using \eqref{eqn:drag} and, in the context of slender-body theory, should be interpreted as the drag-per-unit-span on a 2D cross-section. As seen in the figure, $F_D$ decreases gradually at first, then abruptly tends to zero in the final moments as the vanishes. This behavior, though perhaps surprising, is consistent with the fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. Equation \eqref{eqn:drag} also allows us to calculate the separate contributions of pressure and viscous drag, and we show these with the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. As the body erodes and assumes a more slender form, though, the pressure drag decreases much more rapidly, leaving viscous drag to dominate. This behavior is associated with the tendency towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For the case of a {\em circular} cylinder, the normalization factor of $4 \pi$ produces an {\em exact} relationship (under the slender-body assumptions), where $a$ is taken to be the radius. For other shapes the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale since it generalizes the case of a circle. In Fig.~\ref{area_drag}(b) we show this scaling prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ is the radius of the initially circular body. The scaling law and measured drag agree closely as $t \to t_f$, thus confirming that our simulation accurately captures the slender-body limit as the shape disappears.

@@

With the vanishing rate understood, we now seek to quantify the limiting shape observed in Fig.~\ref{shrink_intface}(a). Certain features of this shape can be predicted analytically by appealing to the principal that, in long time, the shear stress becomes uniformly distributed along the body. In particular, consider the angle formed at the front and rear stagnation points. In a small region around this point, we can neglect body curvature and consider the problem of Stokes flow around an infinite wedge of opening angle $\oangle$. See Fig.~\ref{corner} for a schematic. This problem can be solved exactly in polar coordinates $(r, \theta)$ where $\theta \in [-\thb,\thb]$ and $\thb = \pi - \oangle/2$. (In this section alone, $\theta$ denotes the polar coordinate and not the local tangent angle).

>>>>

	As the body shrinks in accordance with \eqref{area_predict}, the drag that it exerts decreases at a related rate, offering a second check on our numerics. In Fig.~\ref{area_drag}(b), we show the horizontal drag, $F_D$, on the simulated body as it erodes (black curve). The drag is calculated using \eqref{eqn:drag} and, in the context of slender-body theory, this should be interpreted as the drag-per-unit-span. As seen in the figure, $F_D$ decreases gradually at first, then abruptly tends to zero in the final moments as the body vanishes. Though it may seem peculiar, the sudden drop to zero  is consistent with the fact that drag vanishes {\em logarithmically} with body size in 2D Stokes flow. Equation \eqref{eqn:drag} also permits us to calculate the contributions of pressure and viscous drag separately, which are shown by the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. As the body erodes and assumes a more slender form, though, the pressure drag decreases more rapidly, leaving viscous drag to dominate. This behavior is associated with convergence towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For a {\em circular} cylinder, the normalization factor $4 \pi$ produces an {\em exact} relationship (within the context of slender-body theory), where $a$ is taken to be the radius. For other shapes the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale since it generalizes the case of a circle. In Fig.~\ref{area_drag}(b) we show the drag scaling prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ is the radius of the initially circular body. The scaling law and measured drag agree closely as $t \to t_f$, confirming that our simulation accurately captures the slender-body limit as the shape disappears.

@@

With the vanishing rate understood, we now seek to quantify the limiting shape observed in Fig.~\ref{shrink_intface}(a). In particular, consider the corner formed at the front and rear of the body. Locally, we can neglect body curvature and consider the problem of Stokes flow around an infinite wedge; see Fig.~\ref{corner} for a schematic. We seek to predict the wedge's opening angle, $\oangle$, by appealing to the principle that, in long time, erosion uniformly distributes the shear stress along the surface. The surrounding flow can be determined exactly in polar coordinates $(r, \theta)$ where $\theta \in [-\thb,\thb]$ and $\thb = \pi - \oangle/2$. (In this section alone, $\theta$ denotes the polar coordinate and not the local tangent angle).
